# Evaluate Model Performance {#evaluate-model-performance}

Now for the fun part, we get to see the results of our hard work! There are some additional data preparation steps we need to take before we can visualize the results in aggregate; if you are just looking for the charts showing the results they are [shown later on in the "Visualize Results" section below](#visualize-results).

## Summarizing models

<!-- [TODO - HERE add notes on using **postResample()** and explain that if you think about it all you need to evaluate a model is the predictions made (on data never seen by the model) ] -->

Example for one model:

```{r caret_post_resample_example}
postResample(pred = cryptodata_nested$lm_test_predictions[[1]], 
             obs = cryptodata_nested$test_data[[1]]$target_price_24h)
```

We can extract the first element to return the **RMSE** metric, and the second element for the **R Squared (R\^2)** metric. We are using **`[[1]]`** to extract the first element of the [**lm\_test\_predictions**]{style="color: blue;"} and [**test\_data**]{style="color: blue;"} and compare the predictions to the actual value of the [**target\_price24h**]{style="color: blue;"} column.

```{r}
print(paste('Now showing RMSE example:', postResample(pred = cryptodata_nested$lm_test_predictions[[1]], 
                                                      obs = cryptodata_nested$test_data[[1]]$target_price_24h)[[1]]))
```

```{r}
print(paste('Now showing R Squared example:', postResample(pred = cryptodata_nested$lm_test_predictions[[1]], 
                                                           obs = cryptodata_nested$test_data[[1]]$target_price_24h)[[2]]))
```

This model used the earliest subset of the data available for the `r cryptodata_nested$test_data[[1]]$symbol` cryptocurrency. How does the same model used to predict this older subset of the data perform when applied to the most recent subset of the data?

We can get the same summary of results on the **holdout**:

```{r caret_post_resample_example_holdout}
postResample(pred = cryptodata_nested$lm_holdout_predictions[[1]], 
             obs = cryptodata_nested$holdout_data[[1]]$target_price_24h)
```

<!-- If we wrote a [*for loop*](https://r4ds.had.co.nz/iteration.html#for-loops) to iterate through all the options we would get the same results using an **object oriented approach**, this has to do with coding styles and preferences and generally speaking there are many ways of doing the same things in programming. -->

### Comparing Metrics

How do these two results compare and why is this comparison important?

The **RMSE**

<!-- [TODO - ADD HERE explaining RMSE and R^2 metrics and how we will use them to standardize the way we evaluate the effectiveness of a model. We are looking to avoid using models where the results looked good for the test data but then perform much worse than that initial expectation when looking at the holdout results] -->

## Data Prep - Adjust Prices

<!-- [TODO - add note on first doing lm only?] -->

Because cryptocurrencies can vary dramatically in their prices with some trading in the tens of thousands of dollars and others trading for less than a cent, we need to make sure to standardize the RMSE columns to provide a fair comparison for the metric.

Therefore, before using the `postResample()` function, let's convert both the predictions and the target to be the % change in price over the 24 hour period, rather than the change in price (\$).

```{block2, type='infoicon'}
This step is particularly tedious, but it is important. As with the rest of this tutorial, try to understand what we are doing and why even if you find the code overwhelming. All we are doing in this "Adjust Prices" section is we are adjusting all of the prices to be represented as percentage change between observations, which will allow us to draw a fair comparison of the metrics across all cryptocurrencies, which would not be possible using the prices themselves.
```

### Add Last Price

<!-- [TODO - write better] -->

In order to convert the first prediction made to be a percentage, we need to know the previous price.

Function to add last\_price\_train column and append it to the predictions made so we can calculate the % change of the first element, before later removing the value not associated with the predictions:

```{r}
last_train_price <- function(train_data, predictions){
  c(tail(train_data$price_usd,1), predictions)
}
```

Now overwrite the old predictions:

```{r}
cryptodata_nested <- mutate(cryptodata_nested,
                            lm_test_predictions = ifelse(split < 5,
                                                         map2(train_data, lm_test_predictions, last_train_price),
                                                         NA))
```

#### Holdout

Do the same for the holdout. For all of holdout need to take last price point of 5th training split.

```{r}
cryptodata_nested_holdout <- mutate(filter(cryptodata_nested, split == 5),
                                    lm_holdout_predictions = map2(train_data, lm_holdout_predictions, last_train_price))
```

Join the holdout data to all rows based on the cryptocurrency symbol alone:

```{r}
cryptodata_nested <- left_join(cryptodata_nested, 
                               select(cryptodata_nested_holdout, symbol, lm_holdout_predictions),
                               by='symbol')
# Remove unwanted columns
cryptodata_nested <- select(cryptodata_nested, -lm_holdout_predictions.x, -split.y)
# Rename the columns kept
cryptodata_nested <- rename(cryptodata_nested, 
                            lm_holdout_predictions = 'lm_holdout_predictions.y',
                            split = 'split.x')
# Reset the correct grouping structure
cryptodata_nested <- group_by(cryptodata_nested, symbol, split)
```

### Convert to Percentage Change

<!-- [TODO - CLEAN UP EXPLANATIONS HERE!] -->

```{r}
standardize_perc_change <- function(predictions){
  results <- (diff(c(lag(predictions, 1), predictions)) / lag(predictions, 1))*100
  # Exclude the first element, next element will be % change of first prediction
  tail(head(results, length(predictions)), length(predictions)-1)
}
```

<!-- [TODO - Show formula? It's (diff(V1, V2)/abs(V1))*100] -->

Overwrite the old predictions with the new predictions adjusted as a percentage now:

```{r target_adjust_test}
cryptodata_nested <- mutate(cryptodata_nested,
                            lm_test_predictions = ifelse(split < 5,
                                                         map(lm_test_predictions, standardize_perc_change),
                                                         NA),
                            # Holdout for all splits
                            lm_holdout_predictions = map(lm_holdout_predictions, standardize_perc_change))
```

### Actuals

Now do the same thing to the actual prices. Let's make a new column called **actuals** containing the real price values (rather than the predicted ones):

```{r}
actuals_create <- function(train_data, test_data){
  c(tail(train_data$price_usd,1), as.numeric(unlist(select(test_data, price_usd))))
}
```

Use the new function to create the new column `actuals`:

```{r}
cryptodata_nested <- mutate(cryptodata_nested,
                            actuals_test = ifelse(split < 5,
                                             map2(train_data, test_data, actuals_create),
                                             NA))
```

#### Holdout

Again, for the holdout we need the price from the training data of the 5th split

```{r}
cryptodata_nested_holdout <- mutate(filter(cryptodata_nested, split == 5),
                                    actuals_holdout = map2(train_data, holdout_data, actuals_create))
```

Join the holdout data to all rows based on the cryptocurrency symbol alone:

```{r}
cryptodata_nested <- left_join(cryptodata_nested, 
                               select(cryptodata_nested_holdout, symbol, actuals_holdout),
                               by='symbol')
# Remove unwanted columns
cryptodata_nested <- select(cryptodata_nested, -split.y)
# Rename the columns kept
cryptodata_nested <- rename(cryptodata_nested, split = 'split.x')
# Reset the correct grouping structure
cryptodata_nested <- group_by(cryptodata_nested, symbol, split)
```

### Actuals as % Change

Now we can convert the new `actuals` to express the `price_usd` as a % change relative to the previous value using adapting the function from earlier:

```{r}
actuals_perc_change <- function(predictions){
  results <- (diff(c(lag(predictions, 1), predictions)) / lag(predictions, 1))*100
  # Exclude the first element, next element will be % change of first prediction
  tail(head(results, length(predictions)), length(predictions)-1)
}
```

```{r}
cryptodata_nested <- mutate(cryptodata_nested,
                            actuals_test = ifelse(split < 5,
                                             map(actuals_test, actuals_perc_change),
                                             NA),
                            actuals_holdout = map(actuals_holdout, actuals_perc_change))
```

## Review Summary Statistics

Now that we standardized the price to show the percentage change relative to the previous period instead of the price in dollars, we can actually compare the summary statistics across all cryptocurrencies and have it be a fair comparison.

Let's get the same statistic as we did at the beginning of this section, but this time on the standardized values:

```{r}
hydroGOF::rmse(cryptodata_nested$lm_test_predictions[[1]], 
               cryptodata_nested$actuals_test[[1]], 
               na.rm=T)
```

<!-- [TODO - Here add note on using rmse function from hydroGOF package so we can use the na.rm=T argument, would otherwise get NA when a single value is NA. Also explain caveat of that.] -->

### Calculate RMSE

Now make a function to get the RMSE metric for all models:

```{r evaluate_rmse_function}
evaluate_preds_rmse <- function(predictions, actuals){

  hydroGOF::rmse(predictions, actuals, na.rm=T)

}
```

Now we can use map2() to use it to get the RMSE metric for both the test data and the holdout:

```{r add_rmse}
cryptodata_nested <- mutate(cryptodata_nested,
                            lm_rmse_test = unlist(ifelse(split < 5,
                                                         map2(lm_test_predictions, actuals_test, evaluate_preds_rmse),
                                                         NA)),
                            lm_rmse_holdout = unlist(map2(lm_holdout_predictions, actuals_holdout, evaluate_preds_rmse)))
```

<!-- [TODO - Add explanation for ifelse() to add rmse based on test or holdout] -->

Look at the results:

```{r}
select(cryptodata_nested, lm_rmse_test, lm_rmse_holdout)
```

Out of `r nrow(cryptodata_nested)` groups, `r count(ungroup(filter(cryptodata_nested, lm_rmse_test >= lm_rmse_holdout)))$n` had an equal or lower RMSE score for the holdout than the test set.

### Calculate R\^2

Now we can do the same for the R Squared metric:

```{r evaluate_rsquared_function}
evaluate_preds_rsq <- function(predictions, actuals){

  postResample(pred = predictions, obs = actuals)[[2]]

}
```

```{r add_rsq}
cryptodata_nested <- mutate(cryptodata_nested,
                            lm_rsq_test = unlist(ifelse(split < 5,
                                                         map2(lm_test_predictions, actuals_test, evaluate_preds_rsq),
                                                         NA)),
                            lm_rsq_holdout = unlist(map2(lm_holdout_predictions, actuals_holdout, evaluate_preds_rsq)))
```

Look at the results. Wrapping them in `print(n=500)` overwrites the behavior to only give a preview of the data so we can view the full results (up to 500 observations).

```{r}
print(select(cryptodata_nested, lm_rmse_test, lm_rmse_holdout, lm_rsq_test, lm_rsq_holdout), n=500)
```

<!-- WORKING UP TO THIS POINT ON 11/2 -->

## Adjust Prices - All Models

<!-- [TODO - OR IS IT BETTER TO DO ALL AT ONCE ABOVE?] -->

### Add Last Price

And now we can do the same for all the other models

```{r}
cryptodata_nested <- mutate(cryptodata_nested,
                            xgb_test_predictions = ifelse(split < 5,
                                                         map2(train_data, xgb_test_predictions, last_train_price),
                                                         NA),
                            xgbTree_test_predictions = ifelse(split < 5,
                                                         map2(train_data, xgbTree_test_predictions, last_train_price),
                                                         NA),
                            nnet_test_predictions = ifelse(split < 5,
                                                         map2(train_data, nnet_test_predictions, last_train_price),
                                                         NA))
```

##### Holdout

(REMEMBER holdout step for all to get price from 5th train split first)

```{r}
cryptodata_nested_holdout <- mutate(filter(cryptodata_nested, split == 5),
                                    xgb_holdout_predictions = map2(train_data, xgb_holdout_predictions, last_train_price),
                                    xgbTree_holdout_predictions = map2(train_data, xgbTree_holdout_predictions, last_train_price),
                                    nnet_holdout_predictions = map2(train_data, nnet_holdout_predictions, last_train_price))
```

Join the holdout data to all rows based on the cryptocurrency symbol alone:

```{r}
cryptodata_nested <- left_join(cryptodata_nested, 
                               select(cryptodata_nested_holdout, symbol, 
                                      xgb_holdout_predictions, xgbTree_holdout_predictions, nnet_holdout_predictions),
                               by='symbol')
# Remove unwanted columns
cryptodata_nested <- select(cryptodata_nested, -xgb_holdout_predictions.x, 
                            -xgbTree_holdout_predictions.x, -nnet_holdout_predictions.x, -split.y)
# Rename the columns kept
cryptodata_nested <- rename(cryptodata_nested, 
                            xgb_holdout_predictions = 'xgb_holdout_predictions.y',
                            xgbTree_holdout_predictions = 'xgbTree_holdout_predictions.y',
                            nnet_holdout_predictions = 'nnet_holdout_predictions.y',
                            split = 'split.x')
# Reset the correct grouping structure
cryptodata_nested <- group_by(cryptodata_nested, symbol, split)
```

### Convert to % Change

Overwrite the old predictions with the new predictions adjusted as a percentage now:

```{r target_adjust_all_models}
cryptodata_nested <- mutate(cryptodata_nested,
                            xgb_test_predictions = ifelse(split < 5,
                                                         map(xgb_test_predictions, standardize_perc_change),
                                                         NA),
                            # Holdout for all splits
                            xgb_holdout_predictions = map(xgb_holdout_predictions, standardize_perc_change),
                            # xgbTree
                            xgbTree_test_predictions = ifelse(split < 5,
                                                         map(xgbTree_test_predictions, standardize_perc_change),
                                                         NA),
                            # Holdout for all splits
                            xgbTree_holdout_predictions = map(xgbTree_holdout_predictions, standardize_perc_change),
                            # nnet
                            nnet_test_predictions = ifelse(split < 5,
                                                         map(nnet_test_predictions, standardize_perc_change),
                                                         NA),
                            # Holdout for all splits
                            nnet_holdout_predictions = map(nnet_holdout_predictions, standardize_perc_change))
```

### Add Metrics

```{r add_metrics_rest}
cryptodata_nested <- mutate(cryptodata_nested,
                            # XGBoost - RMSE - Test
                            xgb_rmse_test = unlist(ifelse(split < 5,
                                                         map2(xgb_test_predictions, actuals_test, evaluate_preds_rmse),
                                                         NA)),
                            # And holdout:
                            xgb_rmse_holdout = unlist(map2(xgb_holdout_predictions, actuals_holdout ,evaluate_preds_rmse)),
                            # XGBoost - R^2 - Test
                            xgb_rsq_test = unlist(ifelse(split < 5,
                                                         map2(xgb_test_predictions, actuals_test, evaluate_preds_rsq),
                                                         NA)),
                            # And holdout:
                            xgb_rsq_holdout = map2(xgb_holdout_predictions, actuals_holdout, evaluate_preds_rsq),
                            # XGBoost Trees - RMSE - Test
                            xgbTree_rmse_test = unlist(ifelse(split < 5,
                                                         map2(xgbTree_test_predictions, actuals_test, evaluate_preds_rmse),
                                                         NA)),
                            # And holdout:
                            xgbTree_rmse_holdout = unlist(map2(xgbTree_holdout_predictions, actuals_holdout, evaluate_preds_rmse)),
                            # XGBoost Trees - R^2 - Test
                            xgbTree_rsq_test = unlist(ifelse(split < 5,
                                                         map2(xgbTree_test_predictions, actuals_test, evaluate_preds_rsq),
                                                         NA)),
                            # And holdout:
                            xgbTree_rsq_holdout = unlist(map2(xgbTree_holdout_predictions, actuals_holdout, evaluate_preds_rsq)),
                            # Neural Network - RMSE - Test
                            nnet_rmse_test = unlist(ifelse(split < 5,
                                                         map2(nnet_test_predictions, actuals_test, evaluate_preds_rmse),
                                                         NA)),
                            # And holdout:
                            nnet_rmse_holdout = unlist(map2(nnet_holdout_predictions, actuals_holdout, evaluate_preds_rmse)),
                            # Neural Network - R^2 - Test
                            nnet_rsq_test = unlist(ifelse(split < 5,
                                                         map2(nnet_test_predictions, actuals_test, evaluate_preds_rsq),
                                                         NA)),
                            # And holdout:
                            nnet_rsq_holdout = unlist(map2(nnet_holdout_predictions, actuals_holdout, evaluate_preds_rsq)))
```

<!-- Now we have RMSE values for every model created for every cryptocurrency and split of the data: -->

<!-- ```{r} -->

<!-- rmse_scores <- select(cryptodata_nested, lm_rmse_test, xgb_rmse_test, xgbTree_rmse_test, nnet_rmse_test) -->

<!-- # Show RMSE scores -->

<!-- rmse_scores -->

<!-- ``` -->

<!-- And the R Squared values: -->

<!-- ```{r} -->

<!-- rsq_scores <- select(cryptodata_nested, lm_rsq, xgb_rsq, xgbTree_rsq, nnet_rsq) -->

<!-- # Show R^2 scores -->

<!-- rsq_scores -->

<!-- ``` -->

## Eval RMSE

<!-- [TODO - Title?] -->

New object without split column `cryptodata_metrics`:

```{r}
cryptodata_metrics <- group_by(select(ungroup(cryptodata_nested),-split),symbol)
```

### Test

Now for each model we will create a new column giving the average RMSE for the 4 cross-validation test splits:

```{r}
rmse_test <- mutate(cryptodata_metrics,
                      lm = mean(lm_rmse_test, na.rm = T),
                      xgb = mean(xgb_rmse_test, na.rm = T),
                      xgbTree = mean(xgbTree_rmse_test, na.rm = T),
                      nnet = mean(nnet_rmse_test, na.rm = T))
```

Now we can use the `gather()` function to summarize the columns as rows:

```{r gather_rmse}
rmse_test <- unique(gather(select(rmse_test, lm:nnet), 'model', 'rmse', -symbol))
# Show results
rmse_test
```

Now tag the results as having been for the test set

```{r}
rmse_test$eval_set <- 'test'
```

### Holdout

Do the same for the holdout set

```{r}
rmse_holdout <- mutate(cryptodata_metrics,
                      lm = mean(lm_rmse_holdout, na.rm = T),
                      xgb = mean(xgb_rmse_holdout, na.rm = T),
                      xgbTree = mean(xgbTree_rmse_holdout, na.rm = T),
                      nnet = mean(nnet_rmse_holdout, na.rm = T))
```

Now we can use the `gather()` function to summarize the columns as rows:

```{r gather_rmse_holdout}
rmse_holdout <- unique(gather(select(rmse_holdout, lm:nnet), 'model', 'rmse', -symbol))
# Show results
rmse_holdout
```

Now tag the results as having been for the holdout set

```{r}
rmse_holdout$eval_set <- 'holdout'
```

### Union Results

```{r}
rmse_scores <- union(rmse_test, rmse_holdout)
```

## Eval $R^2$

### Test

For each model we will create a new column giving the average R\^2 for the 4 cross-validation test splits:

```{r}
rsq_test <- mutate(cryptodata_metrics,
                      lm = mean(lm_rsq_test, na.rm = T),
                      xgb = mean(xgb_rsq_test, na.rm = T),
                      xgbTree = mean(xgbTree_rsq_test, na.rm = T),
                      nnet = mean(nnet_rsq_test, na.rm = T))
```

Now we can use the `gather()` function to summarize the columns as rows:

```{r gather_rsq}
rsq_test <- unique(gather(select(rsq_test, lm:nnet), 'model', 'rsq', -symbol))
# Show results
rsq_test
```

Now tag the results as having been for the test set

```{r}
rsq_test$eval_set <- 'test'
```

### Holdout

Do the same for the holdout set

```{r}
rsq_holdout <- mutate(cryptodata_metrics,
                      lm = mean(lm_rsq_holdout, na.rm = T),
                      xgb = mean(xgb_rsq_holdout, na.rm = T),
                      xgbTree = mean(xgbTree_rsq_holdout, na.rm = T),
                      nnet = mean(nnet_rsq_holdout, na.rm = T))
```

Now we can use the `gather()` function to summarize the columns as rows:

```{r gather_rsq_holdout}
rsq_holdout <- unique(gather(select(rsq_holdout, lm:nnet), 'model', 'rsq', -symbol))
# Show results
rsq_holdout
```

Now tag the results as having been for the holdout set

```{r}
rsq_holdout$eval_set <- 'holdout'
```

### Union Results

```{r}
rsq_scores <- union(rsq_test, rsq_holdout)
```

## Visualize Results {#visualize-results}

Now we can take the same tools we learned in the [Visualization section from earlier](#visualization) and visualize the results of the models.

### RMSE Visualization

<!-- [TODO - Can I compare test and holdout here?] -->

<!-- ```{r} -->

<!-- ggplot(rmse_scores, aes(x=split, y=rmse, color = model)) + -->

<!--   geom_boxplot() + -->

<!--   geom_point() + -->

<!--   ylim(c(0,50)) + -->

<!--   facet_wrap(~eval_set) -->

<!-- ``` -->

### Both

#### Join Datasets

First join the two

```{r}
plot_scores <- merge(rmse_scores, rsq_scores)
```

#### Plot Results

<!-- [TODO - ADJUST THESE SOME MORE] -->

```{r}
ggplot(plot_scores, aes(x=rsq, y=rmse, color = model)) +
  geom_point() +
  ylim(c(0,10))
```

Running the same code wrapped in the `ggplotly()` function from the `plotly` package [@R-plotly] we can make the chart interactive. Try hovering over the points on the chart with your mouse.

```{r}
ggplotly(ggplot(plot_scores, aes(x=rsq, y=rmse, color = model, symbol = symbol)) +
           geom_point() +
           ylim(c(0,10)),
         tooltip = c("model", "symbol", "rmse", "rsq"))
```

**The additional `tooltip` argument was passed to `ggpltoly()` to specify the label when hovering over the individual points**.

### Results by the Cryptocurrency

We can use the `facet_wrap()` function from `ggplot2` [@R-ggplot2] to create an individual chart for each cryptocurrency:

```{r}
ggplot(plot_scores, aes(x=rsq, y=rmse, color = model)) +
  geom_point() +
  geom_smooth() +
  ylim(c(0,10)) +
  facet_wrap(~symbol)
```

Every 12 hours once this document reaches this point, the results are saved to GitHub using the `pins` package [@R-pins], and a separate script creates the complete dataset over time. You won't be able to run the code shown below (nor do you have a reason to):

```{r upload_results_pin, error=TRUE}
# register board
board_register("github", repo = "predictcrypto/pins", token=pins_key)
# Add current date time
plot_scores$last_refreshed <- Sys.time()
# pin data
pin(plot_scores, board='github', name='crypto_tutorial_results_latest')
```


### Interactive Dashboard
Use the interactive app below to explore the results over time by the individual cryptocurrency. Use the filters on the left sidebar to visualize the results you are interested in:

```{r metrics_by_crypto_shinyapp, out.height = "600px", out.width='100%', echo=F}
knitr::include_app('https://predictcrypto.shinyapps.io/tutorial_latest_model_summary/', height = '600px')
```

*If you have trouble viewing the embedded dashboard you can open it here instead: <https://predictcrypto.shinyapps.io/tutorial_latest_model_summary/>*

The default view shows the holdout results for all models. **Another interesting comparison to make is between the holdout and the test set for fewer models (2 is ideal)**.

```{block2, type='infoicon'}
The app shown above also has a button to **Show Code**. If you were to show the code and copy and paste it into an RStudio session on your computer into a file with the .Rmd file extension and you then *Knit* the file, the same exact app should show up on your computer, no logins or setup outside of the packages required for the code to run; RStudio should automatically prompt you to install packages that are not currently installed on your computer.
```

<!-- TODO - NOT logging MLflow metrics here because needs Python install. Instead run them through R Studio Server. -->

## Visualizations - Historical Metrics

We can pull the same data into this R session using the `pin_get()` function:
```{r}
metrics_historical <- pin_get(name = "full_metrics")
```

The data is limited to metrics for runs from the past 30 days and includes new data every 12 hours. Using the tools we used in the [data prep section](#data-prep), we can answer a couple more questions.

### Best Models

**Overall**, which **model** has the best metrics for all runs from the last 30 days?

#### Summarize the data
```{r summarize_best_models}
# First create grouped data
best_models <- group_by(metrics_historical, model, eval_set)
# Now summarize the data
best_models <- summarize(best_models,
                         rmse = mean(rmse, na.rm=T),
                         rsq  = mean(rsq, na.rm=T))
# Show results
best_models
```

#### Plot RMSE by Model

```{r}
ggplot(best_models, aes(model, rmse, fill = eval_set)) + 
  geom_bar(stat = "identity", position = 'dodge') +
  ggtitle('RMSE by Model', 'Comparing Test and Holdout')
```


<!-- Plotly version: -->
<!-- ```{r plotly_accuracy_over_time} -->
<!-- ggplotly(ggplot(best_models, aes(model, rmse, fill = eval_set)) +  -->
<!--            geom_bar(stat = "identity", position = 'dodge') + -->
<!--            ggtitle('RMSE by Model', 'Comparing Test and Holdout') ) -->
<!-- ``` -->

#### Plot $R^2$ by Model

```{r}
ggplot(best_models, aes(model, rsq, fill = eval_set)) + 
  geom_bar(stat = "identity", position = 'dodge') +
  ggtitle('R^2 by Model', 'Comparing Test and Holdout')
```



### Most Predictable Cryptocurrency


<!-- [TODO - HERE FIND MOST (AND LEAST) PREDICTABLE CRYPTOCURRENCIES] -->

**Overall**, which **cryptocurrency** has the best metrics for all runs from the last 30 days?

#### Summarize the data
```{r summarize_best_cryptos, class.output="scroll-lim"}
# First create grouped data
predictable_cryptos <- group_by(metrics_historical, symbol, eval_set)
# Now summarize the data
predictable_cryptos <- summarize(predictable_cryptos,
                         rmse = mean(rmse, na.rm=T),
                         rsq  = mean(rsq, na.rm=T))
# Arrange from most predictable (according to R^2) to least 
predictable_cryptos <- arrange(predictable_cryptos, desc(rsq))
# Show results
predictable_cryptos
```

Show data using the `formattable` package [@R-formattable] to color code 
```{r}
formattable(predictable_cryptos, list(rmse = color_tile("#71CA97", "red"),
                                     rsq = color_tile("firebrick1", "#71CA97")))
```


### Accuracy Over Time

<!-- [TODO - HERE HAVE ACCURACY OVER TIME. TWO LINES, ONE FOR HOLDOUT AND ONE FOR TEST, AND DO RMSE FIRST AND THEN RSQ] -->

#### Summarize the data
```{r summarize_accuracy_over_time}
# First create grouped data
accuracy_over_time <- group_by(metrics_historical, date_utc)
# Now summarize the data
accuracy_over_time <- summarize(accuracy_over_time, 
                                rmse = mean(rmse, na.rm=T),
                                rsq  = mean(rsq, na.rm=T))
# Ungroup data
accuracy_over_time <- ungroup(accuracy_over_time)
# Show results
accuracy_over_time
```

#### Plot Metrics

```{r plot_accuracy_over_time}
ggplot(accuracy_over_time, aes(x = date_utc, y = rmse, rsq = rsq, group = 1)) +
  # Add both a red point and a red line for RMSE metric
  geom_point(color = 'red', size = 2) +
  geom_line(color = 'red', size = 1) +
  # Also add R^2 to the plot in dark green
  geom_point(aes(x = date_utc, y = rsq), color = 'dark green', size = 2) +
  geom_line(aes(x = date_utc, y = rsq), color = 'dark green', size = 1)
```

<!-- Plotly version: -->
<!-- ```{r plotly_accuracy_over_time} -->
<!-- ggplotly( -->
<!--   ggplot(accuracy_over_time, aes(x = date_utc, y = rmse, rsq = rsq)) + -->
<!--     # Add both a red point and a red line for RMSE metric -->
<!--     geom_point(color = 'red', size = 2) + -->
<!--     geom_line(color = 'red', size = 1) + -->
<!--     # Also add R^2 to the plot in dark green -->
<!--     geom_point(aes(x = date_utc, y = rsq), color = 'dark green', size = 2) + -->
<!--     geom_line(aes(x = date_utc, y = rsq), color = 'dark green', size = 1) -->
<!--   ) -->
<!-- ``` -->

[Refer back to the interactive dashboard](#interactive-dashboard) to take a more specific subset of results instead of the aggregate analysis shown above.


## Making Predictions

Once we have used the information visualized in this section to determine which models we want to go ahead and use, we want to start thinking about how to make predictions on new data.

To make new predictions, we would start by pulling the newest data. Next we want to make sure the data has the exact same columns that were used to train the models minus the target variable which we will predict. We need to apply the exact same data preparation steps that were performed on the data to train the models to the new data in order to make new predictions.

We have omitted these steps to discourage people from performing real trades. Congratulations, you have made it all the way through all of the code! Learn more about why you should not attempt performing trades using the system outlined [in the next section](#considerations).




