# Evaluate Model Performance {#evaluate-model-performance}

Now we get to see the results of our hard work! There are some additional data preparation steps we need to take before we can visualize the results in aggregate; if you are just looking for the charts showing the results they are [shown later on in the "Visualize Results" section below](#interactive-dashboard).

## Summarizing models

Because we know what really happened for the target variable in the test data we used in the previous step, we can get a good idea of how good the model performed on a dataset it has never seen before. We do this to avoid overfitting, which is the idea that the model may work really well on the training data we provided, but not on the new data that we want to predictions on. If the performance on the test set is good, that is a good sign. If the data is split into several subsets and each subset has consistent results for the training and test datasets, that is an even better sign the model may perform as expected.

The first row of the data is for the `r cryptodata_nested$symbol[[1]]` cryptocurrency for the split number `r cryptodata_nested$split[[1]]`. For this row of data (and all others), we made predictions for the [**test_data**]{style="color: blue;"} using a linear regression model and saved the results in the [**lm_test_predictions**]{style="color: blue;"} column. The models were trained on the [**train_data**]{style="color: blue;"} and had not yet seen the results from the [**test_data**]{style="color: blue;"}, so how accurate was the model in its predictions on this data? 

### MAE

Each individual prediction can be compared to the observation of what actually happened, and for each prediction we can calculate the error between the two. We can then take all of the values for the error of the prediction relative to the actual observations, and summarize the performance as a [[**M**ean **A**bsolute **Error**]{style="color: purple;"}](https://en.wikipedia.org/wiki/Mean_absolute_error) (MAE) of those values, which gives us a single value to use as an indicator of the accuracy of the model. The higher the MAE score, the higher the error, meaning the model performs worse when the value is larger.

### RMSE

A common metric to evaluate the performance of a model is the [**R**oot **M**ean **S**quare **E**rror](https://en.wikipedia.org/wiki/Root-mean-square_deviation), which is similar to the MAE but squares and then takes the square root of the values. An interesting implication of this, is that the RMSE will always be larger or equal to the MAE, where a large degree of error on a single observation would get penalized more by the RMSE. The higher the RMSE value, the worse the performance of the model, and can range from 0 to infinity, meaning there is no defined limit on the amount of error you could have (unlike the next metric).

### R Squared

The [**$R^2$**]{style="color: purple;"}, also known as the [**coefficient of determination**]{style="color: purple;"}, is a measure that describes the strength in the correlation between the predictions made and the actual results. A value of 1.0 would mean that the predictions made were exactly identical to the actual results. A perfect score is usually concerning because even a great model shouldn't be exactly 100% accurate and usually indicates a mistake was made that gave away the results to the model and would not perform nearly as good when put into practice in the real world, but in the case of the $R^2$ the higher the score (from 0 to 1) the better.


### Get Metrics

We can return the RMSE and $R^2$ metrics for the `r cryptodata_nested$symbol[[1]]` cryptocurrency and the split number `r cryptodata_nested$split[[1]]` by using the [**postResample()**]{style="color: green;"} function from the [**caret**]{style="color: #ae7b11;"} package:

```{r caret_post_resample_example}
postResample(pred = cryptodata_nested$lm_test_predictions[[1]], 
             obs = cryptodata_nested$test_data[[1]]$target_price_24h)
```

We can extract the first element to return the **RMSE** metric, and the second element for the **R Squared (R\^2)** metric. We are using **`[[1]]`** to extract the first element of the [**lm\_test\_predictions**]{style="color: blue;"} and [**test\_data**]{style="color: blue;"} and compare the predictions to the actual value of the [**target\_price24h**]{style="color: blue;"} column.

This model used the earliest subset of the data available for the `r cryptodata_nested$test_data[[1]]$symbol` cryptocurrency. How does the same model used to predict this older subset of the data perform when applied to the most recent subset of the data from the **holdout**?

We can get the same summary of results comparing the [**lm_holdout_predictions**]{style="color: blue;"} to what actually happened to the [**target_price_24h**]{style="color: blue;"} column of the actual [**holdout_data**]{style="color: blue;"}:

```{r caret_post_resample_example_holdout}
postResample(pred = cryptodata_nested$lm_holdout_predictions[[1]], 
             obs = cryptodata_nested$holdout_data[[1]]$target_price_24h)
```

*The result above may show a value of NA for the RMSE metric. [We will explain and resolve the issue later on](#calculate-rmse-no-NA)*.


### Comparing Metrics

Why not just pick one metric and stick to it? We certainly could, but these two metrics complement each other. For example, if we had a model that always predicts a 0% price change for the time period, the model may have a low error but it wouldn't actually be very informative in the direction or magnitude of those movements and the predictions and actuals would not be very correlated with each other which would lead to a low $R^2$ value. We are using both because it helps paint a more complete picture in this sense, and depending on the task you may want to use a different set of metrics to evaluate the performance. It is also worth mentioning that if your target variable you are predicting is either 0 or 1, this would be a [**classification**]{style="color: purple;"} problem where different metrics become more appropriate to use.

<!-- How do these two results compare and why is this comparison important? The **RMSE** is helpful in understanding the magnitude of the errors of the predictions, while the **$R^2$** helps us determine how easily we can predict future values.  -->

**These are indicators that should be taken with a grain of salt individually, but comparing the results across many different models for the same cryptocurrency can help us determine which models work best for the problem, and then comparing those results across many cryptocurrencies can help us understand which cryptocurrencies we can predict with the most accuracy**.

Before we can draw these comparisons however, we will need to "standardize" the values to create a fair comparison across all dataasets.

<!-- TODO: If for holdout a cryptocurrency averaged less than 1% price movement for the holdout it's not volatile enough for us for it to be relevant, for example if a cryptocurrency is a stablecoin -->


## Data Prep - Adjust Prices

<!-- [TODO - add note on first doing lm only?] -->

Because cryptocurrencies can vary dramatically in their prices with some trading in the tens of thousands of dollars and others trading for less than a cent, we need to make sure to standardize the RMSE columns to provide a fair comparison for the metric.

Therefore, before using the `postResample()` function, let's convert both the predictions and the target to be the % change in price over the 24 hour period, rather than the change in price (\$).

```{block2, type='infoicon'}
This step is particularly tedious, but it is important. As with the rest of this tutorial, try to understand what we are doing and why even if you find the code overwhelming. All we are doing in this "Adjust Prices" section is we are adjusting all of the prices to be represented as percentage change between observations, which will allow us to draw a fair comparison of the metrics across all cryptocurrencies, which would not be possible using the prices themselves. If you want to [skip the tedious steps and want to see the performance of the models visualized, click here to skip ahead.](#visualize-results)
```

### Add Last Price

<!-- [TODO - write better - KEEP GOING 11/14] -->

In order to convert the first prediction made to be a percentage, we need to know the previous price, which would be the last observation from the train data. Therefore, let's make a function to add the [**last_price_train**]{style="color: blue;"} column and append it to the predictions made so we can calculate the % change of the first element relative to the last observation in the train data, before later removing the value not associated with the predictions:

```{r}
last_train_price <- function(train_data, predictions){
  c(tail(train_data$price_usd,1), predictions)
}
```

**We will first perform all steps on the linear regression models to make the code a little more digestible, and we will then perform the [same steps for the rest of the models](#adjust-prices-all-models).**

#### Test 

Overwrite the old predictions for the first 4 splits of the test data using the new function created above:

```{r}
cryptodata_nested <- mutate(cryptodata_nested,
                            lm_test_predictions = ifelse(split < 5,
                                                         map2(train_data, lm_test_predictions, last_train_price),
                                                         NA))
```
*The [**mutate()**]{style="color: green;"} function is used to create the new column [**lm_test_predictions**]{style="color: blue;"} assigning the value only for the first 4 splits where the test data would actually exist (the 5th being the holdout set) using the [**ifelse()**]{style="color: green;"} function*.


#### Holdout

Do the same but for the holdout now. For the holdout we need to take the last price point of the 5th split:
```{r}
cryptodata_nested_holdout <- mutate(filter(cryptodata_nested, split == 5),
                                    lm_holdout_predictions = map2(train_data, lm_holdout_predictions, last_train_price))
```

Now join the holdout data to all rows based on the cryptocurrency symbol alone:

```{r}
cryptodata_nested <- left_join(cryptodata_nested, 
                               select(cryptodata_nested_holdout, symbol, lm_holdout_predictions),
                               by='symbol')
# Remove unwanted columns
cryptodata_nested <- select(cryptodata_nested, -lm_holdout_predictions.x, -split.y)
# Rename the columns kept
cryptodata_nested <- rename(cryptodata_nested, 
                            lm_holdout_predictions = 'lm_holdout_predictions.y',
                            split = 'split.x')
# Reset the correct grouping structure
cryptodata_nested <- group_by(cryptodata_nested, symbol, split)
```

### Convert to Percentage Change

Now we have everything we need to accurately calculate the percentage change between observations including the first one. Let's make a new function to calculate the percentage change:

```{r}
standardize_perc_change <- function(predictions){
  results <- (diff(c(lag(predictions, 1), predictions)) / lag(predictions, 1))*100
  # Exclude the first element, next element will be % change of first prediction
  tail(head(results, length(predictions)), length(predictions)-1)
}
```

Overwrite the old predictions with the new predictions adjusted as a percentage now:

```{r target_adjust_test}
cryptodata_nested <- mutate(cryptodata_nested,
                            lm_test_predictions = ifelse(split < 5,
                                                         map(lm_test_predictions, standardize_perc_change),
                                                         NA),
                            # Holdout for all splits
                            lm_holdout_predictions = map(lm_holdout_predictions, standardize_perc_change))
```

### Actuals

Now do the same thing to the actual prices. Let's make a new column called [**actuals**]{style="color: blue;"} containing the real price values (rather than the predicted ones):

```{r}
actuals_create <- function(train_data, test_data){
  c(tail(train_data$price_usd,1), as.numeric(unlist(select(test_data, price_usd))))
}
```

Use the new function to create the new column [**actuals**]{style="color: blue;"}:

```{r}
cryptodata_nested <- mutate(cryptodata_nested,
                            actuals_test = ifelse(split < 5,
                                             map2(train_data, test_data, actuals_create),
                                             NA))
```

#### Holdout

Again, for the holdout we need the price from the training data of the 5th split to perform the first calculation:
```{r}
cryptodata_nested_holdout <- mutate(filter(cryptodata_nested, split == 5),
                                    actuals_holdout = map2(train_data, holdout_data, actuals_create))
```

Join the holdout data to all rows based on the cryptocurrency symbol alone:

```{r}
cryptodata_nested <- left_join(cryptodata_nested, 
                               select(cryptodata_nested_holdout, symbol, actuals_holdout),
                               by='symbol')
# Remove unwanted columns
cryptodata_nested <- select(cryptodata_nested, -split.y)
# Rename the columns kept
cryptodata_nested <- rename(cryptodata_nested, split = 'split.x')
# Reset the correct grouping structure
cryptodata_nested <- group_by(cryptodata_nested, symbol, split)
```


### Actuals as % Change

Now we can convert the new [**actuals**]{style="color: blue;"} to express the [**price_usd**]{style="color: blue;"} as a % change relative to the previous value using adapting the function from earlier:

```{r}
actuals_perc_change <- function(predictions){
  results <- (diff(c(lag(predictions, 1), predictions)) / lag(predictions, 1))*100
  # Exclude the first element, next element will be % change of first prediction
  tail(head(results, length(predictions)), length(predictions)-1)
}
```

```{r}
cryptodata_nested <- mutate(cryptodata_nested,
                            actuals_test = ifelse(split < 5,
                                             map(actuals_test, actuals_perc_change),
                                             NA),
                            actuals_holdout = map(actuals_holdout, actuals_perc_change))
```

## Review Summary Statistics

Now that we standardized the price to show the percentage change relative to the previous period instead of the price in dollars, we can actually compare the summary statistics across all cryptocurrencies and have it be a fair comparison.

Let's get the same statistic as we did at the beginning of this section, but this time on the standardized values. This time to calculate the RMSE error metric let's use the [**rmse()**]{style="color: green;"} function from the [**hydroGOF**]{style="color: #ae7b11;"} package because it allows us to set the [**`na.rm = T`**]{style="color: blue;"} parameter, and otherwise one NA value would return NA for the overall RMSE:

```{r}
hydroGOF::rmse(cryptodata_nested$lm_test_predictions[[1]], 
               cryptodata_nested$actuals_test[[1]], 
               na.rm=T)
```


### Calculate R\^2

Now we can do the same for the R Squared metric using the same [**postResample()**]{style="color: green;"} function that we used at [the start of this section](#evaluate-model-performance):

```{r evaluate_rsquared_function}
evaluate_preds_rsq <- function(predictions, actuals){

  postResample(pred = predictions, obs = actuals)[[2]]

}
```

```{r add_rsq}
cryptodata_nested <- mutate(cryptodata_nested,
                            lm_rsq_test = unlist(ifelse(split < 5,
                                                         map2(lm_test_predictions, actuals_test, evaluate_preds_rsq),
                                                         NA)),
                            lm_rsq_holdout = unlist(map2(lm_holdout_predictions, actuals_holdout, evaluate_preds_rsq)))
```

Look at the results:

```{r}
select(cryptodata_nested, lm_rsq_test, lm_rsq_holdout)
```


### Calculate RMSE {#calculate-rmse-no-NA}

Similarly let's make a function to get the RMSE metric for all models:

```{r evaluate_rmse_function}
evaluate_preds_rmse <- function(predictions, actuals){

  hydroGOF::rmse(predictions, actuals, na.rm=T)

}
```

Now we can use the [**map2()**]{style="color: green;"} function to use it to get the RMSE metric for both the test data and the holdout:

```{r add_rmse}
cryptodata_nested <- mutate(cryptodata_nested,
                            lm_rmse_test = unlist(ifelse(split < 5,
                                                         map2(lm_test_predictions, actuals_test, evaluate_preds_rmse),
                                                         NA)),
                            lm_rmse_holdout = unlist(map2(lm_holdout_predictions, actuals_holdout, evaluate_preds_rmse)))
```

<!-- [TODO - Add explanation for ifelse() to add rmse based on test or holdout] -->

Look at the results. Wrapping them in [**`print(n=500)`**]{style="color: blue;"} overwrites the behavior to only give a preview of the data so we can view the full results (up to 500 observations).

```{r}
print(select(cryptodata_nested, lm_rmse_test, lm_rmse_holdout, lm_rsq_test, lm_rsq_holdout), n=500)
```

Out of `r nrow(cryptodata_nested)` groups, `r count(ungroup(filter(cryptodata_nested, lm_rmse_test >= lm_rmse_holdout)))$n` had an equal or lower RMSE score for the holdout than the test set.


## Adjust Prices - All Models{#adjust-prices-all-models}

Let's repeat the same steps that we outlined above for all models.

### Add Last Price

```{r}
cryptodata_nested <- mutate(cryptodata_nested,
                            # XGBoost:
                            xgb_test_predictions = ifelse(split < 5,
                                                         map2(train_data, xgb_test_predictions, last_train_price),
                                                         NA),
                            # Neural Network:
                            nnet_test_predictions = ifelse(split < 5,
                                                         map2(train_data, nnet_test_predictions, last_train_price),
                                                         NA),
                            # Random Forest:
                            rf_test_predictions = ifelse(split < 5,
                                                         map2(train_data, rf_test_predictions, last_train_price),
                                                         NA),
                            # PCR:
                            pcr_test_predictions = ifelse(split < 5,
                                                         map2(train_data, pcr_test_predictions, last_train_price),
                                                         NA))
```

##### Holdout

```{r}
cryptodata_nested_holdout <- mutate(filter(cryptodata_nested, split == 5),
                                    # XGBoost:
                                    xgb_holdout_predictions = map2(train_data, xgb_holdout_predictions, last_train_price),
                                    # Neural Network:
                                    nnet_holdout_predictions = map2(train_data, nnet_holdout_predictions, last_train_price),
                                    # Random Forest:
                                    rf_holdout_predictions = map2(train_data, rf_holdout_predictions, last_train_price),
                                    # PCR:
                                    pcr_holdout_predictions = map2(train_data, pcr_holdout_predictions, last_train_price))
```

Join the holdout data to all rows based on the cryptocurrency symbol alone:

```{r}
cryptodata_nested <- left_join(cryptodata_nested, 
                               select(cryptodata_nested_holdout, symbol, 
                                      xgb_holdout_predictions, nnet_holdout_predictions, 
                                      rf_holdout_predictions, pcr_holdout_predictions),
                               by='symbol')
# Remove unwanted columns
cryptodata_nested <- select(cryptodata_nested, -xgb_holdout_predictions.x, 
                            -nnet_holdout_predictions.x,-rf_holdout_predictions.x, 
                            -pcr_holdout_predictions.x, -split.y)
# Rename the columns kept
cryptodata_nested <- rename(cryptodata_nested, 
                            xgb_holdout_predictions = 'xgb_holdout_predictions.y',
                            nnet_holdout_predictions = 'nnet_holdout_predictions.y',
                            rf_holdout_predictions = 'rf_holdout_predictions.y',
                            pcr_holdout_predictions = 'pcr_holdout_predictions.y',
                            split = 'split.x')
# Reset the correct grouping structure
cryptodata_nested <- group_by(cryptodata_nested, symbol, split)
```

### Convert to % Change

Overwrite the old predictions with the new predictions adjusted as a percentage now:

```{r target_adjust_all_models}
cryptodata_nested <- mutate(cryptodata_nested,
                            # XGBoost:
                            xgb_test_predictions = ifelse(split < 5,
                                                         map(xgb_test_predictions, standardize_perc_change),
                                                         NA),
                            # holdout - all splits
                            xgb_holdout_predictions = map(xgb_holdout_predictions, standardize_perc_change),
                            # nnet:
                            nnet_test_predictions = ifelse(split < 5,
                                                         map(nnet_test_predictions, standardize_perc_change),
                                                         NA),
                            # holdout - all splits
                            nnet_holdout_predictions = map(nnet_holdout_predictions, standardize_perc_change),
                            # Random Forest:
                            rf_test_predictions = ifelse(split < 5,
                                                         map(rf_test_predictions, standardize_perc_change),
                                                         NA),
                            # holdout - all splits
                            rf_holdout_predictions = map(rf_holdout_predictions, standardize_perc_change),
                            # PCR:
                            pcr_test_predictions = ifelse(split < 5,
                                                         map(pcr_test_predictions, standardize_perc_change),
                                                         NA),
                            # holdout - all splits
                            pcr_holdout_predictions = map(pcr_holdout_predictions, standardize_perc_change))
```

### Add Metrics

Add the RMSE and $R^2$ metrics:
```{r add_metrics_rest}
cryptodata_nested <- mutate(cryptodata_nested,
                            # XGBoost - RMSE - Test
                            xgb_rmse_test = unlist(ifelse(split < 5,
                                                         map2(xgb_test_predictions, actuals_test, evaluate_preds_rmse),
                                                         NA)),
                            # And holdout:
                            xgb_rmse_holdout = unlist(map2(xgb_holdout_predictions, actuals_holdout ,evaluate_preds_rmse)),
                            # XGBoost - R^2 - Test
                            xgb_rsq_test = unlist(ifelse(split < 5,
                                                         map2(xgb_test_predictions, actuals_test, evaluate_preds_rsq),
                                                         NA)),
                            # And holdout:
                            xgb_rsq_holdout = unlist(map2(xgb_holdout_predictions, actuals_holdout, evaluate_preds_rsq)),
                            # Neural Network - RMSE - Test
                            nnet_rmse_test = unlist(ifelse(split < 5,
                                                         map2(nnet_test_predictions, actuals_test, evaluate_preds_rmse),
                                                         NA)),
                            # And holdout:
                            nnet_rmse_holdout = unlist(map2(nnet_holdout_predictions, actuals_holdout, evaluate_preds_rmse)),
                            # Neural Network - R^2 - Test
                            nnet_rsq_test = unlist(ifelse(split < 5,
                                                         map2(nnet_test_predictions, actuals_test, evaluate_preds_rsq),
                                                         NA)),
                            # And holdout:
                            nnet_rsq_holdout = unlist(map2(nnet_holdout_predictions, actuals_holdout, evaluate_preds_rsq)),
                            # Random Forest - RMSE - Test
                            rf_rmse_test = unlist(ifelse(split < 5,
                                                         map2(rf_test_predictions, actuals_test, evaluate_preds_rmse),
                                                         NA)),
                            # And holdout:
                            rf_rmse_holdout = unlist(map2(rf_holdout_predictions, actuals_holdout, evaluate_preds_rmse)),
                            # Random Forest - R^2 - Test
                            rf_rsq_test = unlist(ifelse(split < 5,
                                                         map2(rf_test_predictions, actuals_test, evaluate_preds_rsq),
                                                         NA)),
                            # And holdout:
                            rf_rsq_holdout = unlist(map2(rf_holdout_predictions, actuals_holdout, evaluate_preds_rsq)),
                            # PCR - RMSE - Test
                            pcr_rmse_test = unlist(ifelse(split < 5,
                                                         map2(pcr_test_predictions, actuals_test, evaluate_preds_rmse),
                                                         NA)),
                            # And holdout:
                            pcr_rmse_holdout = unlist(map2(pcr_holdout_predictions, actuals_holdout, evaluate_preds_rmse)),
                            # PCR - R^2 - Test
                            pcr_rsq_test = unlist(ifelse(split < 5,
                                                         map2(pcr_test_predictions, actuals_test, evaluate_preds_rsq),
                                                         NA)),
                            # And holdout:
                            pcr_rsq_holdout = unlist(map2(pcr_holdout_predictions, actuals_holdout, evaluate_preds_rsq)))
```

Now we have RMSE and $R^2$ values for every model created for every cryptocurrency and split of the data:

```{r}
select(cryptodata_nested, lm_rmse_test, lm_rsq_test, lm_rmse_holdout, lm_rsq_holdout)
```
*Only the results for the linear regression model are shown. There are equivalent columns for the XGBoost, neural network, random forest and PCR models.*


## Evaluate Metrics Across Splits

Next, let's evaluate the metrics across all [**splits**]{style="color: blue;"} and keeping moving along with the [model validation plan as was originally outlined](#time-aware-cross-validation). Let's create a new dataset called [**cryptodata_metrics**][**splits**]{style="color: blue;"} that is not grouped by the [**split**]{style="color: blue;"} column and is instead only grouped by the [**symbol**]{style="color: blue;"}:

```{r}
cryptodata_metrics <- group_by(select(ungroup(cryptodata_nested),-split),symbol)
```

### Evaluate RMSE Test

Now for each model we can create a new column giving the average RMSE for the 4 cross-validation test splits:

```{r}
rmse_test <- mutate(cryptodata_metrics,
                      lm = mean(lm_rmse_test, na.rm = T),
                      xgb = mean(xgb_rmse_test, na.rm = T),
                      nnet = mean(nnet_rmse_test, na.rm = T),
                      rf = mean(rf_rmse_test, na.rm = T),
                      pcr = mean(pcr_rmse_test, na.rm = T))
```

Now we can use the [**gather()**]{style="color: green;"} function to summarize the columns as rows:

```{r gather_rmse}
rmse_test <- unique(gather(select(rmse_test, lm:pcr), 'model', 'rmse', -symbol))
# Show results
rmse_test
```

Now tag the results as having been for the test set:

```{r}
rmse_test$eval_set <- 'test'
```

### Holdout

Now repeat the same process for the holdout set:

```{r}
rmse_holdout <- mutate(cryptodata_metrics,
                      lm = mean(lm_rmse_holdout, na.rm = T),
                      xgb = mean(xgb_rmse_holdout, na.rm = T),
                      nnet = mean(nnet_rmse_holdout, na.rm = T),
                      rf = mean(rf_rmse_holdout, na.rm = T),
                      pcr = mean(pcr_rmse_holdout, na.rm = T))
```

Again, use the [**gather()**]{style="color: green;"} function to summarize the columns as rows:

```{r gather_rmse_holdout}
rmse_holdout <- unique(gather(select(rmse_holdout, lm:pcr), 'model', 'rmse', -symbol))
# Show results
rmse_holdout
```

Now tag the results as having been for the holdout set:

```{r}
rmse_holdout$eval_set <- 'holdout'
```

### Union Results

Now we can [**union()**]{style="color: green;"} the results to stack the rows from the two datasets on top of each other:
```{r}
rmse_scores <- union(rmse_test, rmse_holdout)
```

## Evaluate R^2

Now let's repeat the same steps we took for the RMSE metrics above for the $R^2$ metric as well.

### Test

For each model again we will create a new column giving the average $R^2$ for the 4 cross-validation test splits:

```{r}
rsq_test <- mutate(cryptodata_metrics,
                      lm = mean(lm_rsq_test, na.rm = T),
                      xgb = mean(xgb_rsq_test, na.rm = T),
                      nnet = mean(nnet_rsq_test, na.rm = T),
                      rf = mean(rf_rsq_test, na.rm = T),
                      pcr = mean(pcr_rsq_test, na.rm = T))
```

Now we can use the [**gather()**]{style="color: green;"} function to summarize the columns as rows:

```{r gather_rsq}
rsq_test <- unique(gather(select(rsq_test, lm:pcr), 'model', 'rsq', -symbol))
# Show results
rsq_test
```

Now tag the results as having been for the test set

```{r}
rsq_test$eval_set <- 'test'
```

### Holdout

Do the same and calculate the averages for the holdout sets:

```{r}
rsq_holdout <- mutate(cryptodata_metrics,
                      lm = mean(lm_rsq_holdout, na.rm = T),
                      xgb = mean(xgb_rsq_holdout, na.rm = T),
                      nnet = mean(nnet_rsq_holdout, na.rm = T),
                      rf = mean(rf_rsq_holdout, na.rm = T),
                      pcr = mean(pcr_rsq_holdout, na.rm = T))
```

Now we can use the [**gather()**]{style="color: green;"} function to summarize the columns as rows:

```{r gather_rsq_holdout}
rsq_holdout <- unique(gather(select(rsq_holdout, lm:pcr), 'model', 'rsq', -symbol))
# Show results
rsq_holdout
```

Now tag the results as having been for the holdout set:

```{r}
rsq_holdout$eval_set <- 'holdout'
```

### Union Results

```{r}
rsq_scores <- union(rsq_test, rsq_holdout)
```

## Visualize Results {#visualize-results}

Now we can take the same tools we learned in the [Visualization section from earlier](#visualization) and visualize the results of the models.

### RMSE Visualization

<!-- [TODO - Can I compare test and holdout here?] -->

<!-- ```{r} -->

<!-- ggplot(rmse_scores, aes(x=split, y=rmse, color = model)) + -->

<!--   geom_boxplot() + -->

<!--   geom_point() + -->

<!--   ylim(c(0,50)) + -->

<!--   facet_wrap(~eval_set) -->

<!-- ``` -->

### Both

Now we have everything we need to use the two metrics to compare the results.

#### Join Datasets

First join the two objects [**rmse_scores**]{style="color: blue;"} and [**rsq_scores**]{style="color: blue;"} into the new object [**plot_scores]{style="color: blue;"}:

```{r}
plot_scores <- merge(rmse_scores, rsq_scores)
```

#### Plot Results

Now we can plot the results on a chart:

```{r}
ggplot(plot_scores, aes(x=rsq, y=rmse, color = model)) +
  geom_point() +
  ylim(c(0,10))
```

Running the same code wrapped in the [**ggplotly()**]{style="color: green;"} function from the [**plotly**]{style="color: #ae7b11;"} package (as we [have already done](#plotly)) we can make the chart interactive. Try hovering over the points on the chart with your mouse.

```{r}
ggplotly(ggplot(plot_scores, aes(x=rsq, y=rmse, color = model, symbol = symbol)) +
           geom_point() +
           ylim(c(0,10)),
         tooltip = c("model", "symbol", "rmse", "rsq"))
```

**The additional [**tooltip**]{style="color: blue;"} argument was passed to [**ggpltoly()**]{style="color: green;"} to specify the label when hovering over the individual points**.

### Results by the Cryptocurrency

We can use the [**facet_wrap()**]{style="color: green;"} function from [**ggplot2**]{style="color: #ae7b11;"} to create an individual chart for each cryptocurrency:

```{r}
ggplot(plot_scores, aes(x=rsq, y=rmse, color = model)) +
  geom_point() +
  geom_smooth() +
  ylim(c(0,10)) +
  facet_wrap(~symbol)
```

Every 12 hours once this document reaches this point, the results are saved to GitHub using the [**pins**]{style="color: #ae7b11;"} package (which we [used to read in the data at the start](#pull-the-data)), and a separate script running on a different server creates the complete dataset in our database over time. You won't be able to run the code shown below (nor do you have a reason to):

```{r upload_results_pin, error=TRUE}
# register board
board_register("github", repo = "predictcrypto/pins", token=pins_key)
# Add current date time
plot_scores$last_refreshed <- Sys.time()
# pin data
pin(plot_scores, board='github', name='crypto_tutorial_results_latest')
```

## Interactive Dashboard
Use the interactive app below to explore the results over time by the individual cryptocurrency. Use the filters on the left sidebar to visualize the results you are interested in:

```{r metrics_by_crypto_shinyapp, out.height = "600px", out.width='100%', echo=F}
knitr::include_app('https://predictcrypto.shinyapps.io/tutorial_latest_model_summary/', height = '600px')
```

*If you have trouble viewing the embedded dashboard you can open it here instead: <https://predictcrypto.shinyapps.io/tutorial_latest_model_summary/>*

The default view shows the holdout results for all models. **Another interesting comparison to make is between the holdout and the test set for fewer models (2 is ideal)**.

```{block2, type='infoicon'}
The app shown above also has a button to **Show Code**. If you were to show the code and copy and paste it into an RStudio session on your computer into a file with the .Rmd file extension and you then *Knit* the file, the same exact app should show up on your computer, no logins or setup outside of the packages required for the code to run; RStudio should automatically prompt you to install packages that are not currently installed on your computer.
```

<!-- TODO - NOT logging MLflow metrics here because needs Python install. Instead run them through R Studio Server. -->

## Visualizations - Historical Metrics

We can pull the same data into this R session using the `pin_get()` function:
```{r}
metrics_historical <- pin_get(name = "full_metrics")
```

The data is limited to metrics for runs from the past 30 days and includes new data every 12 hours. Using the tools we used in the [data prep section](#data-prep), we can answer a couple more questions.

### Best Models

**Overall**, which **model** has the best metrics for all runs from the last 30 days?

#### Summarize the data
```{r summarize_best_models}
# First create grouped data
best_models <- group_by(metrics_historical, model, eval_set)
# Now summarize the data
best_models <- summarize(best_models,
                         rmse = mean(rmse, na.rm=T),
                         rsq  = mean(rsq, na.rm=T))
# Show results
best_models
```

#### Plot RMSE by Model

```{r}
ggplot(best_models, aes(model, rmse, fill = eval_set)) + 
  geom_bar(stat = "identity", position = 'dodge') +
  ggtitle('RMSE by Model', 'Comparing Test and Holdout')
```


<!-- Plotly version: -->
<!-- ```{r plotly_accuracy_over_time} -->
<!-- ggplotly(ggplot(best_models, aes(model, rmse, fill = eval_set)) +  -->
<!--            geom_bar(stat = "identity", position = 'dodge') + -->
<!--            ggtitle('RMSE by Model', 'Comparing Test and Holdout') ) -->
<!-- ``` -->

#### Plot $R^2$ by Model

```{r}
ggplot(best_models, aes(model, rsq, fill = eval_set)) + 
  geom_bar(stat = "identity", position = 'dodge') +
  ggtitle('R^2 by Model', 'Comparing Test and Holdout')
```



### Most Predictable Cryptocurrency


<!-- [TODO - HERE FIND MOST (AND LEAST) PREDICTABLE CRYPTOCURRENCIES] -->

**Overall**, which **cryptocurrency** has the best metrics for all runs from the last 30 days?

#### Summarize the data
```{r summarize_best_cryptos, class.output="scroll-lim"}
# First create grouped data
predictable_cryptos <- group_by(metrics_historical, symbol, eval_set)
# Now summarize the data
predictable_cryptos <- summarize(predictable_cryptos,
                         rmse = mean(rmse, na.rm=T),
                         rsq  = mean(rsq, na.rm=T))
# Arrange from most predictable (according to R^2) to least 
predictable_cryptos <- arrange(predictable_cryptos, desc(rsq))
# Show results
predictable_cryptos
```

```{css, echo=FALSE}
.scroll-100 {
  max-height: 100px;
  overflow-y: auto;
  background-color: inherit;
}
```

Show the top 15 most predictable cryptocurrencies (according to the $R^2$) using the `formattable` package [@R-formattable] to color code the cells:
```{r, class.output="scroll-100"}
formattable(head(predictable_cryptos ,15), 
            list(rmse = color_tile("#71CA97", "red"),
                 rsq = color_tile("firebrick1", "#71CA97")))
```


### Accuracy Over Time

<!-- [TODO - HERE HAVE ACCURACY OVER TIME. TWO LINES, ONE FOR HOLDOUT AND ONE FOR TEST, AND DO RMSE FIRST AND THEN RSQ] -->

#### Summarize the data
```{r summarize_accuracy_over_time}
# First create grouped data
accuracy_over_time <- group_by(metrics_historical, date_utc)
# Now summarize the data
accuracy_over_time <- summarize(accuracy_over_time, 
                                rmse = mean(rmse, na.rm=T),
                                rsq  = mean(rsq, na.rm=T))
# Ungroup data
accuracy_over_time <- ungroup(accuracy_over_time)
# Show results
accuracy_over_time
```

#### Plot RMSE

<!-- [TODO - here is it better to facet_wrap by eval_set to compare test/holdout?] -->

Remember, for RMSE the lower the score, the more accurate the models were.

```{r plot_accuracy_over_time_rmse}
ggplot(accuracy_over_time, aes(x = date_utc, y = rmse, group = 1)) +
  # Plot RMSE over time
  geom_point(color = 'red', size = 2) +
  geom_line(color = 'red', size = 1)
```

#### Plot R^2

For the R^2 recall that we are looking at the correlation between the predictions made and what actually happened, so the higher the score the better, with a maximum score of 1 that would mean the predictions were 100% correlated with each other and therefore identical.

```{r plot_accuracy_over_time_rsq}
ggplot(accuracy_over_time, aes(x = date_utc, y = rsq, group = 1)) +
  # Plot R^2 over time
  geom_point(aes(x = date_utc, y = rsq), color = 'dark green', size = 2) +
  geom_line(aes(x = date_utc, y = rsq), color = 'dark green', size = 1)
```


[Refer back to the interactive dashboard](#interactive-dashboard) to take a more specific subset of results instead of the aggregate analysis shown above.


<!-- ## Making Predictions -->

<!-- Once we have used the information visualized in this section to determine which models we want to go ahead and use, we want to start thinking about how to make predictions on new data. -->

<!-- To make new predictions, we would start by pulling the newest data. Next we want to make sure the data has the exact same columns that were used to train the models minus the target variable which we will predict. We need to apply the exact same data preparation steps that were performed on the data to train the models to the new data in order to make new predictions. -->

<!-- We have omitted these steps to discourage people from performing real trades. Congratulations, you have made it all the way through all of the code! Learn more about why you should not attempt performing trades using the system outlined [in the next section](#considerations). -->



<!-- [TODO - HERE unnest data for holdout predictions and write the data to pins! From there pick up using RStudio Server script] -->

<!-- HIDDEN: Add new pin with actual predictions from holdout for true comparisons! -->
```{r record_holdout_predictions, include=F, error=T}
record_holdout_preds_cbind <- function(holdout_data, holdout_predictions){
  left_join(mutate(as.data.frame(holdout_predictions), num = row_number()), 
            mutate(cryptodata_nested$holdout_data[[1]], num = row_number()))
}

# Use new function
cryptodata_nested <- mutate(cryptodata_nested,
                            lm_holdout_predictions = map2(holdout_data, lm_holdout_predictions, record_holdout_preds_cbind),
                            xgb_holdout_predictions = map2(holdout_data, xgb_holdout_predictions, record_holdout_preds_cbind),
                            nnet_holdout_predictions = map2(holdout_data, nnet_holdout_predictions, record_holdout_preds_cbind),
                            rf_holdout_predictions = map2(holdout_data, rf_holdout_predictions, record_holdout_preds_cbind),
                            pcr_holdout_predictions = map2(holdout_data, pcr_holdout_predictions, record_holdout_preds_cbind))

# UNNEST the data:

# lm:
lm_holdout_preds <- select(mutate(unnest(mutate(cryptodata_nested,
                                                lm_holdout_predictions = map2(holdout_data, lm_holdout_predictions,
                                                                              record_holdout_preds_cbind)), lm_holdout_predictions),
                                  model = 'lm'), holdout_predictions:target_price_24h, -num)
# xgb:
xgb_holdout_preds <- select(mutate(unnest(mutate(cryptodata_nested,
                                                xgb_holdout_predictions = map2(holdout_data, xgb_holdout_predictions,
                                                                              record_holdout_preds_cbind)), xgb_holdout_predictions),
                                  model = 'xgb'), holdout_predictions:target_price_24h, -num)
# nnet:
nnet_holdout_preds <- select(mutate(unnest(mutate(cryptodata_nested,
                                                nnet_holdout_predictions = map2(holdout_data, nnet_holdout_predictions,
                                                                              record_holdout_preds_cbind)), nnet_holdout_predictions),
                                  model = 'nnet'), holdout_predictions:target_price_24h, -num)
# random forest:
rf_holdout_preds <- select(mutate(unnest(mutate(cryptodata_nested,
                                                rf_holdout_predictions = map2(holdout_data, rf_holdout_predictions,
                                                                              record_holdout_preds_cbind)), rf_holdout_predictions),
                                  model = 'rf'), holdout_predictions:target_price_24h, -num)
# PCR:
pcr_holdout_preds <- select(mutate(unnest(mutate(cryptodata_nested,
                                                pcr_holdout_predictions = map2(holdout_data, pcr_holdout_predictions,
                                                                              record_holdout_preds_cbind)), pcr_holdout_predictions),
                                  model = 'pcr'), holdout_predictions:target_price_24h, -num)

# Now union all data together:
full_holdout_preds <- dplyr::bind_rows(lm_holdout_preds, xgb_holdout_preds, nnet_holdout_preds, rf_holdout_preds, pcr_holdout_preds)

# Only Keep data where we don't yet know what the price was 24 hours later:
write_holdout_preds <- filter(full_holdout_preds, is.na(target_price_24h))

# Before overwriting pins data, pull it so I can add new targets
old_holdout_preds_pins <- pin_get('crypto_tutorial_holdout_predictions_old')

# PIN NEW DATA so it can be read by RStudio Server process

# register board
board_register("github", repo = "predictcrypto/pins", token=pins_key)
# Add current date time
write_holdout_preds$last_refreshed <- Sys.time()
# pin data
pin(write_holdout_preds, board='github', name='crypto_tutorial_holdout_predictions_old')

# new data pull
prices_predictions <- left_join(write_holdout_preds, old_holdout_preds_pins)
# pin data with new info on prices
pin(write_holdout_preds, board='github', name='crypto_tutorial_holdout_predictions_n')
```
